Visualizing Conditional Relationships

### 2017-08-31 paper/main.tex
new abstract

\title{Visualizing Conditional Relationships}
\author{Hannah Aizenman}
\committee{Dr. Michael Grossberg(Advisor), Dr. Robert Haralick, Dr. Huy Vo}
\submitted{}
\abstract{The context surrounding data is often as important as the
  measurements themselves. Income data provides a lens into demographic
  inequalities when separated by race and gender; temperature data indicates
  seasons when plotted by time. These are relatively simple relationships, but
 many datasets are embedded with rather complex conditional relationships. This
 survey gives an overview of techniques showing conditional dependencies
 amongst distributions and densities. 
}
\begin{document}
\makefrontmatter

@@

\subfile{sections/distdist.tex}
%%\subfile{sections/distdens.tex}
%%\subfile{sections/densdist.tex}
%%\subfile{sections/densdens.tex}

>>>>

\title{Visualizing Conditional Relationships}
\author{Hannah Aizenman}
\committee{Dr. Michael Grossberg(Advisor), Dr. Robert Haralick, Dr. Huy Vo}
\submitted{}
\abstract{Many datasets have qualitative and quantitative variables and  also
  often have functional attributes like time and location and visualizations
  provide views on a subset of these variables and how they relate to each
  other. For example a timeseries plot will keep the link between the variable
  and time but lose the information encoded in the space variable. These
  relationships can be modeled as conditional probabilities  such as P(V|t),
  and visualizations can be organized based on which of these probabilities are
  implicitly preserved in the visualization techniques. Modern visualizations
  in particular try to tackle the difficulties of displaying these functional
  relationships as the number of variables and the number of functional
  dependencies (for example lat, lon, height, time, and type) increase. }
\begin{document}
\makefrontmatter

@@

\subfile{sections/distdist.tex}
\subfile{sections/distdens.tex}
\subfile{sections/densdist.tex}
\subfile{sections/densdens.tex}

### 2017-08-31 paper/sections/densdist.tex
new abstract

\documentclass[../main.text]{subfiles}
\begin{document}
\section{Densities from Distributions}

@@

\begin{figure}
  \includegraphics{flexaxis.png}
  \caption{Figure taken from \cite{claessen_flexible_2011} illustrates the following common
    methods for visualizing multivariate datasets: matrix of scatters, parallel coordinates and radar charts.
  \label{fig:flexlink}
\end{figure}

@@

\begin{figure}
  \includegraphics{parallelsets.png}
  \caption{On the left is the categorical view, where users can define
    categories based on other categorical information. For example, in this
    view working family with children is defined as a household with one or
    more children. On the right is a parallel coordinate graph where the
    dominant category informs the colors, the size of the box is relative to
    the category size, and the histograms show the frequencies of those
    categories relative to the connecting categories (market type and income
    respectively)}
  \label{fig:parallelsets}
\end{figure}

@@

\begin{figure}
  \includegraphics{corr.png}
  \label{caption} Zhang et al create an initial map (a) that they then apply an
  edge filter to (b). They then allow the user to filter out a variable and
  either highlight only the vertexes of that variable (c) or the edges incident
  on that variable(d).
  \label{fig:corr}
\end{figure}

Often though it's important to retain the internal structure of the
quantitative data, such as when the data is a function of time and
space. Zhang et al propose a method that retains data and allows for
visualizing more than just pairwise relationships\cite{zhang_visual_2015}. They
introduce a concept of a correlation map, as shown in
figure~\ref{fig:corr}. The authors build out a graph where every variable is a
vertex and the edges are correlations between the connecting vertexes. The
vertexes are sized relative to the variance in the data. The edges are colored
red for negative correlations and green for positive correlations. To convert
the categorical data to numerical space, they developed an algorithm where when
mapped to numbers, the residual sum of squares of a logistic regression of the
categorical variables  would be minimized. They then filter edges with low
correlations and on variables of interest.
\end{document}

>>>>

\documentclass[../main.text]{subfiles}
\begin{document}
\section{Quantitative Densities Conditioned on Qualitative Distributions}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{flexaxis.png}
  \caption{This figure taken from a paper on improving multivariate data visualization using flexible linked axis \cite{claessen_flexible_2011} shows the common techniques of matrix of scatters, parallel coordinates and radar charts.}
  \label{fig:flexlink}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{parallelsets.png}
  \caption{On the left is the categorical view, where users can define
    categories based on other categorical information. For example, in this
    view working family with children is defined as a household with one or
    more children. On the right is a parallel coordinate graph where the
    dominant category informs the colors, the size of the box is relative to
    the category size, and the histograms show the frequencies of those
    categories relative to the connecting categories (market type and income
    respectively)}
  \label{fig:parallelsets}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{corr.png}
  \caption{Zhang et al create an initial map (a) that they then apply an
  edge filter to (b). They then allow the user to filter out a variable and
  either highlight only the vertexes of that variable (c) or the edges incident
  on that variable(d).}
  \label{fig:corr}
\end{figure}

Often though it's important to retain the internal structure of the
quantitative data, such as when the data is a function of time and
space. Zhang et al propose a method that retains data and allows for
visualizing more than just pairwise relationships \cite{zhang_visual_2015}. They
introduce a concept of a correlation map, as shown in
figure~\ref{fig:corr}. The authors build out a graph where every variable is a
vertex and the edges are correlations between the connecting vertexes. The
vertexes are sized relative to the variance in the data. The edges are colored
red for negative correlations and green for positive correlations. To convert
the categorical data to numerical space, they developed an algorithm where when
mapped to numbers, the residual sum of squares of a logistic regression of the
categorical variables  would be minimized. They then filter edges with low
correlations and on variables of interest.
\end{document}

### 2017-08-31 paper/sections/distdist.tex
new abstract

\section{Distributions}

@@

 %%\begin{figure}
  %% \includegraphics[width=\textwidth]{upsetfig1}
   %%\caption{UpSet dashboard displaying the relationships amongst movie
  %%genres. On the left is the set visualization; the columns are the sets and
    %% the rows are the exclusive intersections or aggregates. On the right is
    %% the element view, and the scatter plot is a comparison of two sets with
     %%each dot equaling one element in the set.}
   %%\label{fig:upsetfig}
 %%\end{figure}

>>>>

\section{Qualitative Variables Conditioned on Qualitative Distributions}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{upsetfig1}
   \caption{UpSet dashboard displaying the relationships amongst movie genres. On the left is the set visualization; the columns are the sets and
   the rows are the exclusive intersections or aggregates. On the right is the element view, and the scatter plot is a comparison of two sets with each dot equaling one element in the set.}
   \label{fig:upsetfig}
\end{figure}
